### 2018-05-11 Paper.tex
updates

\documentclass[12pt,oneside]{amsart}
%\usepackage{tweaklist}
\usepackage{cancel}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage{cuted}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{afterpage}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr. Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
%,pagebackref=true
\usepackage{tabu} %requires array.

\def\file{\mathcal{F}}
\def\filehash{\sigma_F}
\def\key{K}
\def\keyhash{\sigma_K}
\def\hash{sha3}
\def\alice{\mathcal{A}}
\def\bob{\mathcal{B}}
\def\ether{\xi}
\def\zksnark{\zeta}
\def\encrypt{\varepsilon}
\def\decrypt{\delta}
\def\crp{\textit{Content Release Problem} }
\def\olp{\textit{Ownership Ledger Problem} }
\def\nb{\textbf{NB:} }
\newcommand{\st}{\mbox{ s.t. }}
\newcommand{\wrt}{\mbox{ w.r.t. }}
\newcommand{\etal}{\mbox{ et al. }}

%This should be the last package before \input{Version.tex}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
% "hyperref loads the url package internally. Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref. This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options. Try loading the url package earlier than the package that requires it. If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.

\makeatletter
\newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
\makeatother
\usepackage{seqsplit}

% For formatting
%\usepackage{underscore}
%\usepackage{lipsum} % to generate filler text for testing of document rendering
\usepackage[english]{babel}
\usepackage[autostyle]{csquotes}
\MakeOuterQuote{"}

\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142

\input{Version.tex}
% Default rendering options
\definecolor{pagecolor}{rgb}{1,0.98,0.9}
\def\YellowPaperVersionNumber{unknown revision}
\IfFileExists{Options.tex}{\input{Options.tex}}

\newcommand{\hcancel}[1]{%
    \tikz[baseline=(tocancel.base)]{
        \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
        \draw[black] (tocancel.south west) -- (tocancel.north east);
    }%
}%

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\newcommand*\eg{e.g.\@\xspace}
\newcommand*\Eg{e.g.\@\xspace}
\newcommand*\ie{i.e.\@\xspace}
%\renewcommand{\itemhook}{\setlength{\topsep}{0pt}  \setlength{\itemsep}{0pt}\setlength{\leftmargin}{15pt}}

% REPORT REQUIREMENTS
% -------------------
%   6-8 pages
%
% Introduction
%   Problem Statement
%   Motivation
% Results
%   Solution Architecture
%   Implementation Details with Fuctionalities
% Summary/Conclusion
% Code Documentation

\title{ML-Runner: A Scalable Cloud System for Machine \& Distributed Learning \\ {\smaller \textbf{Project Report \YellowPaperVersionNumber}}}
\author{
    Barak Ben Noon (bb2762@columbia.edu) \\
    James Maxwell Stenger (jms2431@columbia.edu)
}
\begin{document}
\pagecolor{pagecolor}

\begin{abstract}
Machine Learning (ML) models are playing an ever increasing role in consumer and enterprise applications.
Since roughly 2012, when Alex Krizhevsky \etal won the ImageNet Large-Scale Visual Recognition Challenge with a deep convolutional neural network algorithm called SuperVision,
machine learning systems have improved exponentially.
Today, ML systems now rival human capabilities in a large variety of application domains.
Asynchronous Deep Reinforcement Learning models such as OpenAI's \href{https://github.com/openai/retro}{Retro} defeat professional Dota2 players;
ML models power smart-assistant software such as Apple's Siri, Amazon's Alexa, and Google Home;
and in general ML is used effectively in applications ranging from stock trading to social network analysis.

ML models often include a number of hyperparameters which must be tuned in order to effectively train the model.
This tuning process is generally performed by data scientists who must be careful to avoid over or underfitting the model so as to optimize its performance on data outside the training set.
This process of adjusting the hyperparameters involves training and evaluating the model with each ``configuration'' of the parameters.

We sought to design an online platform which would allow end users to programmatically specify a set of ``configurations'' for an ML model which they want to tune.
The platform subsequently trains and evaluates the model in each those of configurations in the cloud, and makes the results available to the end user.
\end{abstract}

\maketitle
\setlength{\columnsep}{20pt}
\begin{multicols}{2}

% -------------- INTRODUCTION --------------
% ------------------------------------------
\section{Introduction}\label{sec:introduction}
In short, the Ethereum network provides us with a framework for creating distributed application that can act as a trustless, uncensorable digital content store.

\subsection{Problem Statement} \label{ch:problem_statement}
\subsection{Motivation} \label{ch:motivation}
bla bla \textit{Such an open system would resolve, or atleast mitigate, the spoofed reviews issue.}

\subsection{Related Work} \label{ch:previous}
\href{https://comet.ml/}{Comet.ml} and \href{https://neptune.ml/}{Nuptune.ml} are two web services which facilitate the running of machine learning experiments with a web frontend that allows for experiment tracking and hyperparameter optimization.
They leave the actual training and testing of the ML experiments to the end user, \ie the end user must configure his computing resources to train and test the model himself.
Our service seeks to simplify the actual running of these experiments, and could potentially integrate with these services in the future.

% -------------- RESULTS --------------
% -------------------------------------
\section{Results} \label{ch:results}
\subsection{Solution Architecture}
The general architecture for our application was very similar to the one used for the three class assignments throughout the semester,
\ie a web frontend written in React that connected to a variety of microservices through AWS API Gateway.

We hosted the frontend on an AWS EC2 instance running an nginx server, but of course it could alternatively be hosted on S3 in the fashion described in the class assignments.
Authentication for API gateway was handled using AWS IAM and Cognito, which we integrated with the frontend by using the react-cognito node package.

API gateway routed to three lambda functions which used a variety of other AWS services:

\begin{itemize}
  \item SQS -- Used to manage the runs of each experiment
  \item S3 -- Saved the docker containers we generated for each configuration of the models
  \item Elastic Beanstalk -- Used to train and evaluate each model configuration as present in its respective docker container
  \item Cloudwatch -- Used to schedule lambda functions which periodically pulled runs from SQS and executed them on EB
\end{itemize}

\subsection{Implementation Details} \label{ch:solidity_work}

\end{multicols}
\subsubsection{Frontend}
We implemented a React front end which we deployed on an Nginx server on an AWS EC2 instance.
Our front end made use of a number of frameworks including:

\begin{itemize}
  \item aws-sdk
  \item react
  \item react-cognito (for user management and authentication with AWS Cognito)
  \item react-redux
  \item react-router
  \item react-router-redux
\end{itemize}

We provided pages for logging in, registering, learning about the service, checking one's profile, viewing a dashboard, creating new experiments, and viewing and running experiments.

\subsubsection{Backend: LF1}
See \ref{code:lf1} for the full javascript code.

\subsubsection{Backend: LF2}
See \ref{code:lf2} for the full javascript code.

\subsubsection{Backend: LF3}
See \ref{code:lf3} for the full javascript code.

\begin{multicols}{2}
% -------------- FUTURE DIRECTIONS --------------
% -----------------------------------------------
\section{Future Directions} \label{ch:future}
In our application we used AWS Elastic Beanstalk to run the containerized experiments for each configuration, since this is what we were familiar with.
It may be preferable to use AWS Elastic Container Service instead, since we already package the code in docker containers, though due to time constraints we did not get a chance to investigate this option.

% -------------- CONCLUSION --------------
% ----------------------------------------
\section{Conclusion} \label{ch:conclusion}
We have presented in this paper an architecture and near-complete implementation of an online service for concurrently training and testing machine learning models with a variety of hyperparameter configurations.

\section{Acknowledgements}
This paper was produced as part of Sambit Sahu's Spring 2018
\textit{Cloud Computing and Big Data} Course at Columbia University (\url{http://www.cs.columbia.edu/~sambits/})

\section{Availability}
The source of this paper is maintained at \url{https://github.com/jmstenger/ml-runner-paper/}.
All project code can be found under the 6998 Organization at \url{https://github.com/6998}.
See \ref{ch:code} for a walkthrough of the code.

\end{multicols}
\bibliographystyle{plainnat}
\bibliography{Biblio}

\appendix

% -------------- CODE DOCUMENTATION --------------
% ------------------------------------------------
\section{Code Documentation} \label{ch:code}
\subsection{LF1: create-experiments} \label{code:lf1}
The javascript code for LF1 is shown below; the full deployment package is available at \url{https://github.com/6998/LF1}.
The lambda code is contained there as `index.js'.

\begin{Verbatim}[fontsize=\tiny, frame=single]
const fs = require('fs');
const path = require('path');
const AWS = require('aws-sdk');
const async = require('async');
const dynamoose = require('dynamoose');
const randomstring = require("randomstring");

const s3 = new AWS.S3({apiVersion: '2006-03-01'});
const sqs = new AWS.SQS({apiVersion: '2012-11-05'});

const DIR = process.env.DIR

// docker template
let dockerfile = "FROM python:3.4-alpine\n" +
"ADD . /code\n" +
"WORKDIR /code\n" +
"EXPOSE 5000\n" +
"RUN pip install -r requirements.txt\n" +
"CMD [\"python\", \"app.py\"]\n"

let finalArr = []; // for the combinations

// zipping part
var file_system = require('fs');
var archiver = require('archiver');

/* Dynamo */
const Run = dynamoose.model('Run', {
  id: {
    type: String,
    hashKey: false,
  },
  options: Object,
  s3: Object,
  err: Object,
  user: String,
  projectName: String,
  projectId: String,
  mode: String
}, { update: false });

const saveRuns = (items, user, cb) => {
  const arr = Object.keys(items).map(function(key) {
    return items[key];
  });
  Run.batchPut(arr, function (err, p) {
    if (err) {
      console.log(err);
    } else {
    }
    return cb(err);
  })
}

const createArray = (item) => {
  if (item.rangeOrType === "range") {
    const tmp = [];
    const from = parseFloat(item.range.from);
    const to = parseFloat(item.range.to);
    const jumps = parseFloat(item.jumps);
    for (let value = from; value <= to; value += jumps) {
      tmp.push({
        value: value.toFixed(3),
        key: item.key
      });
    }
    return tmp;
  } else if (item.rangeOrType === "options") {
    const tmp = item.values.split(",");
    return tmp.map(el => {
      return {
        value: el,
        key: item.key
      }
    })
  } else {
    return [];
  }
};

const allPossibleCombos = (options) => {
  let optionsArr = [];
  for (let k in options) { // create an array for each options
    optionsArr.push(options[k]);
  }

  optionsArr = optionsArr.map(el=>{ // convert each option to a range
    return createArray(el)
  });

  optionsArr = optionsArr.sort((a,b)=>{ // sort by length bigger first
    return b.length-a.length
  });

  optionsArr[0].forEach(item=>{
    finalArr.push([item]);
  });

  optionsArr = optionsArr.slice(1); // remove the first line

  optionsArr.forEach((opArr,a)=>{ // for each array of option
    const newSets = [];
    opArr.forEach((el,b)=>{ // for each element in the array
      const tmp = finalArr.map(el=>{
        return el.map(el2=>Object.assign(el2, {}));
      }); // deep clone

      tmp.forEach((arr,c)=>{ // for each set in the final array
        arr.push(el);
      }); // end adding element for each sub-set

      newSets.push(tmp);
    }); // end each element in the the range

    finalArr = [];
    newSets.forEach((item,d)=>{
      finalArr = finalArr.concat(item)
    })
  });

  return finalArr;
};

const renderFile = (codeArr, options)=>{
  options.forEach(item => {
    const key = item.key.split("_");
    codeArr[key[0]][key[1]] = `RUNNER_${item.key}`
  });

  let finalFile = "";
  codeArr.forEach(line=>{
    line.forEach(word=>{
      finalFile += word.replace(/\&nbsp\;/g, "\t")
    });
    finalFile += "\n";
  });

  return {finalFile, options};
};

const createZipAndUpload = (fileObj, jobNum, cb)=>{
  const file = fileObj.finalFile;
  const options = fileObj.options;
  // inject the os and the:

  let split = file.split("\n");

  // find the last line with the import statment
  let lastImportLine = 0;
  split.forEach((el, line)=>{
    if(el.includes("import")) { lastImportLine = line; }
  });

  const envVarsArr = options.map(item=>{
    return `RUNNER_${item.key}=os.environ['RUNNER_${item.key}']`;
  });

  const envVarsArrForDocker = options.map(item=>{
    return `ENV RUNNER_${item.key}=os.environ['RUNNER_${item.key}']`;
  });

  // create the app.py
  const part1 = split.slice(0, lastImportLine+1).join("\n");
  const part2 = envVarsArr.join("\n");
  const part3 = split.slice(lastImportLine+1, split.length).join("\n");
  const finalFile = [].concat(part1).concat(part2).concat(part3).join("\n");
  dockerfile = dockerfile += envVarsArrForDocker.join("\n");

  //create the Dockerfile
  fs.writeFileSync(`${DIR}/Dockerfile`, dockerfile);
  fs.writeFileSync(`${DIR}/app.py`, finalFile);

  // save the file
  const rand = randomstring.generate(7);
  const fileName = `job_num_${jobNum}_rand_${rand}.zip`;
  const filePath = `${DIR}/${fileName}`
  const output = file_system.createWriteStream(filePath);
  const archive = archiver('zip');

  output.on('close', function () {
    console.log('archiver has been finalized and the output file descriptor has closed.');
  });

  archive.on('error', function(err){
    console.log("error saving file:", err)
    // throw err;
  });

  output.on('end', function() {
    console.log('Data has been drained');
  });

  output.on('finish', function() {
    console.log('finish');
    const uploadParams = {Bucket: "docker-6998", Key: '', Body: ''};
    const fileStream = fs.createReadStream(filePath);

    fileStream.on('error', function(err) {
      console.log('fileStream Error', err);
    });
    uploadParams.Body = fileStream;
    uploadParams.Key = path.basename(fileName);

    // call S3 to retrieve upload file to specified bucket
    s3.putObject(uploadParams, function (err, data) {
      if (err) {
        console.log("putObject Error", err);
      } if (data) {
      }
      cb(err, data, uploadParams.Key, fileName);
    });
  });

  archive.pipe(output);
  archive.file(`${DIR}/requirements.txt`, "")
  archive.file(`${DIR}/Dockerfile`, "")
  archive.file(`${DIR}/app.py`, "")
  archive.file(`${DIR}/docker-compose.yml`, "")
  archive.finalize();
};

const addToSQS = (user, projectId, projectName, key, cb)=>{
  const MessageAttributes = {
    User: {
      DataType: "String",
      StringValue: user
    },
    ProjectName: {
      DataType: "String",
      StringValue: projectName
    },
    ProjectId: {
      DataType: "String",
      StringValue: projectId
    },
    S3key: {
      DataType: "String",
      StringValue: key
    },
  };

  var params = {
    DelaySeconds: 1,
    MessageAttributes,
    MessageBody: "New Version",
    QueueUrl: "https://sqs.us-east-1.amazonaws.com/906385631751/ml-runner"
  };

  sqs.sendMessage(params, function(err, data) {
    if (err) {
      console.log("sendMessage with sqs:", err);
    } else {
    }
    cb(err)
  });
}

exports.handler = (event, context, callback) => {
  let options = event.options;
  let codeArr = event.codeArr;
  let user = event.user;
  let projectName = event.projectName;

  const projectId = randomstring.generate(7);
  const allCombos = allPossibleCombos(options);

  let i = 0;
  const re = {};

  async.eachSeries(allCombos, (op, cb) => {
    const file = renderFile(codeArr, op);
    createZipAndUpload(file, i++, (err, data, key, fileName) => {
      const id = `job_${i}`;
      re[id] = {
        file: file.finalFile,
        options: file.options,
        s3: data,
        err,
        id: fileName,
        user: user,
        projectName,
        projectId,
        mode: "init"
      };
      addToSQS(user, projectId, projectName, key, (err)=>{
        cb();
      })
    })
  }, (err1) => {
    saveRuns(re, user, (err) => {
      console.log("save err", err);
      return callback(null, {files: re});
    })
  });
};

exports.createArray = createArray;
exports.allPossibleCombos = allPossibleCombos;
\end{Verbatim}

\subsection{LF2: create-EB} \label{code:lf2}
The javascript code for LF2 is shown below; the full deployment package is available at \url{https://github.com/6998/LF2}.
The lambda code is contained there as `index.js'.

\begin{Verbatim}[fontsize=\tiny, frame=single]
const AWS = require('aws-sdk');
const sqs = new AWS.SQS({apiVersion: '2012-11-05'});
const async = require('async');
const fs = require('fs');
const path = require('path');
const s3 = new AWS.S3({apiVersion: '2006-03-01'});
const randomstring = require("randomstring");
const dynamoose = require('dynamoose');
const elasticbeanstalk = new AWS.ElasticBeanstalk();

exports.handler = (event, context, callback) => {
  var queueURL = "https://sqs.us-east-1.amazonaws.com/906385631751/ml-runner";

  var params = {
    AttributeNames: [
      "SentTimestamp"
    ],
    MaxNumberOfMessages: 10,
    MessageAttributeNames: [
      "All"
    ],
    QueueUrl: queueURL,
    VisibilityTimeout: 0,
    WaitTimeSeconds: 0
  };

  sqs.receiveMessage(params, function(err, data) {
    async.eachSeries(data.Messages, (item, cb)=>{
      console.log(tem.MessageAttributes)

      const projectName = item.MessageAttributes.ProjectName.StringValue;
      const ProjectId = item.MessageAttributes.ProjectId.StringValue;
      const S3key = item.MessageAttributes.S3key.StringValue;
      const name = `runner_${projectName}_${ProjectId}`;

      if(item.Body === "New Version") {
        const EBparams = {
          ApplicationName: name,
          AutoCreateApplication: true,
          Description: `runner_${projectName}`,
          Process: true,
          SourceBundle: {
            S3Bucket: "docker-6998",
            S3Key: S3key
          },
          VersionLabel: S3key
        };

        var deleteParams = {
          QueueUrl: queueURL,
          ReceiptHandle: item.ReceiptHandle
        };

        sqs.deleteMessage(deleteParams, function(err, data) {
            console.log("Delete Error", err);
        });

        // Create EB version
        elasticbeanstalk.createApplicationVersion(EBparams, function(err, data) {
          console.log("elasticbeanstalk ", err);
          cb();
        });
      } else if(item.Body === "Run") {
        const rand = randomstring.generate(3);
        let env = S3key.split["."];
        env = env[0];

        var params = {
          ApplicationName: name,
          CNAMEPrefix: projectName+rand,
          EnvironmentName: env,
          SolutionStackName: "64bit Amazon Linux 2017.09 v2.9.2 running Docker 17.12.0-ce",
          VersionLabel: S3key
        };

        // Launch EB instance
        elasticbeanstalk.createEnvironment(params, function(err, data) {
          console.log("createEnvironment: ", err);
          cb();
        });
      } else {
        cb();
      }

    }, ()=>{
      return callback(null, true);
    })
  });
};
\end{Verbatim}

\subsection{LF3: create-run} \label{code:lf3}
The javascript code for LF3 is shown below; the full deployment package is available at \url{https://github.com/6998/LF3}.
The lambda code is contained there as `index.js'.

\begin{Verbatim}[fontsize=\tiny, frame=single]
  a
\end{Verbatim}

In \ref{ch:future}, we outline how expensive on-chain storage is.

% -------------- TERMINOLOGY --------------
% -----------------------------------------
\section{Terminology} \label{ch:Terminology}

\begin{description}
\item[address] triple abc
\end{description}

\appendix
\end{document}